---
title: "Curve Fitting Plot1 inset"
output:
  pdf_document: default
  html_notebook: default
---
```{r load data, include=FALSE}
library(ggplot2)
library(gridExtra)
load("year_counts.RData")
```

In this notebook we explore the claim that the distribution of patent numbers over time follows a power law distribution. 

As we saw in the reproducing valverde notebook the valverde suggests that the cumluative number of patents is a power law $N(t) \approx t^\theta$. From our fit and including more modern data we can see that this fit doesn't seem appropriate. The two plots below show the cdf of patents each year on a linear and log-log scale respectively. 

```{r plot1_inset, fig.height=4, fig.width=11, echo=F}
gg1b <- ggplot(data = year_counts, aes(x = rel_year, y = cum_count)) + 
    geom_point() +
    theme_bw() +
    labs(x = "t", y = "N(t)")

grid.arrange(gg1b, gg1b + scale_x_log10() + scale_y_log10(), ncol=2)
```

By transforming the data in different ways we can fit standard linear models to predict power law and exponential functions respectively:
The exponential function clearly performs better with low p value and higher R squared. 
```{r, warning=FALSE, echo=F}
attach(year_counts, warn.conflicts = FALSE)
exp.model <- lm(log(count) ~ rel_year)
summary(exp.model)

pl.model <- lm(log(count) ~ log(rel_year))
summary(pl.model)
```

We can visualise these fits onto the data. This clearly shows how the exponential function fits but more importantly when translated to non-cdf how the exponential function still nicely fits the data but the power law looks very odd and is only fitting to one cluster of data, it almost looks like a textbook example of high bias. 
```{r, fig.height=4, fig.width=11, echo=F}
exp.pred <- exp(predict(exp.model, interval = "confidence", level = 0.95))
pl.pred <- exp(predict(pl.model, newdata = data.frame(rel_year = 1:40), interval = "confidence", level = 0.95))

gg1 <- ggplot(data = year_counts, aes(x = rel_year, y = count)) + 
    geom_point() + 
    geom_line(aes(y = exp.pred[,"fit"])) +
    geom_ribbon(aes(ymax = exp.pred[,"upr"], ymin = exp.pred[,"lwr"]), alpha = 0.3) +
    theme_bw() + 
    geom_line(aes(y = pl.pred[,"fit"], x = 1:40), linetype = "dashed")

gg <- ggplot(data = year_counts, aes(x = rel_year, y = cum_count)) +
    geom_point() +
    geom_line(aes(y = cumsum(exp.pred[,"fit"]))) +
    geom_ribbon(aes(ymax = cumsum(exp.pred[,"upr"]), ymin = cumsum(exp.pred[,"lwr"])), alpha = 0.3) +
    theme_bw() +
    geom_line(aes(y = cumsum(pl.pred[,"fit"])), linetype = "dashed")

grid.arrange(gg1, gg, gg + scale_x_log10() + scale_y_log10(), ncol = 3)
```

# Questions
 - Should I try to fit other distributions
 - Is this enough analysis of this or should I do something more thorough? What would I do?
 - Plots In report should I have plots like this? If so I will make them 
```{r save, include=F}
png("Figures/patentCountFit.png"); gg1; dev.off()
png("Figures/patentCountFit_cum.png"); gg; dev.off()
png("Figures/patentCountFit_cum_loglog.png"); gg + scale_x_log10() + scale_y_log10(); dev.off()
```

